Why stemmed points? Understanding the usage of the Korea Late Paleolithic stemmed points using tip cross-sectional area

### 2022-03-16 analysis/paper/paper.Rmd
editing rmd file (plots and text)

---
title: "Title Goes Here"
author:
  - Myfirstname Mylastname:
      email: fl@oneplace.edu
      institute: [UofO]
      correspondence: true
  - Yourfirstname Yourlastname:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

tcsa <- mydata %>%
  mutate(min_tcsa = 0.5 *Width * Thickness) %>%
  mutate(max_tcsa = 0.5 *Length * Thickness) %>%
  #omit very large TCSA, could be measurement error
  filter(min_tcsa <= 250) %>%
  # group rare raw materials
  mutate(raw_materials_top = fct_lump_min(Raw_material, 10)) %>%
  mutate(site_name_top = fct_lump_min(Site_name, 5))
```

```{r}
# all artefacts overall
ggplot (tcsa,
        aes(1, min_tcsa)) +
  # show regions for tools from Table 3 of Lomabrd 2021
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
    annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 4,
           ymax = 18,
           fill = "grey80") +
      annotate("text", 
               label = "Poisoned arrowheads",
           x = 1.3,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 18,
           ymax = 53,
           fill = "grey50") +
      annotate("text", 
               label = "Arrowheads",
           x = 1.3,
           y = (53-13) /2 + 13) +
      annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 53,
           ymax = 76,
           fill = "grey90") +
      annotate("text", 
               label = "Dart tips",
           x = 1.3,
           y = (76-53) /2 + 53 ) +
      annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 76,
           ymax = 98,
           fill = "grey70") +
      annotate("text", 
               label = "Javelin",
           x = 1.3,
           y = (98-76) /2 + 76 ) +
     annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 98,
           ymax = 230,
           fill = "grey95") +
      annotate("text", 
               label = "Thrusting spear",
           x = 1.3,
           y = (230-98) /2 + 98 ) +
  geom_boxplot(alpha = 0.1) +
  geom_quasirandom(alpha = 0.4,
                   size = 3) +

  theme_bw()
```

```{r}
#### histogram for min tcsa to prepare for exploring the data
all <- ggplot(tcsa,
              aes(min_tcsa)) +
  geom_histogram() +
  theme_bw()
```

@@

```{r}
# Facet by raw materials
all +  facet_wrap(~raw_materials_top,
                  scales = "free_y",
                  ncol = 1)
```

```{r}
# remove NAs
tcsa_mass_no_na <-
  tcsa %>%
 drop_na(`Weight(g)`)

# how does the size-tcsa relationship vary by raw material?
ggplot(tcsa_mass_no_na,
       aes(`Weight(g)`,
           min_tcsa,
           colour = raw_materials_top)) +
  geom_point() +
  theme_bw()
```

@@

```{r}
# Facet by site
all +  facet_wrap(~site_name_top,
                 scales = "free_y",
                  ncol = 1)
```

>>>>

---
title: "Why stemmed points? Understanding the usage of the Korea Late Paleolithic stemmed points using tip cross-sectional area"
author:
  - Gayoung Park: 
      email: gayoungp@uw.edu
      institute: [Uof W]
      correspondence: true
  - Marlize Lombard:
      email: mlombard@uj.ac.za
      institute: [UofJ]
      correspondence: false
  - Ben Marwick:
      email: bmarwick@uw.edu
      institute: [UofW]
      correspondence: false
institute:
  - UofW: University of Washington
  - UofJ: Palaeo-Research Institute, University of Johannesburg

date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

tcsa <- mydata %>%
  mutate(min_tcsa = 0.5 *Width * Thickness) %>%
  mutate(max_tcsa = 0.5 *Length * Thickness) %>%
  #omit very large TCSA, could be measurement error
  filter(min_tcsa <= 250) %>%
  # group rare raw materials
  mutate(raw_materials_top = fct_lump_min(Raw_material, 10)) %>%
  mutate(site_name_top = fct_lump_min(Site_name, 5))

```

```{r, fig.width= 4.5, fig.height=6}
# all artefacts overall
ggplot (tcsa,
        aes(1, min_tcsa)) +
  # show regions for tools from Table 3 of Lomabrd 2021
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
    annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = 3,
               label = "Poisoned arrowheads",
           x = 1.3,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 13,
           ymax = 53,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = 3,
               label = "Arrowheads",
           x = 1.3,
           y = (53-13) /2 + 13) +
      annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = 3,
               label = "Dart tips",
           x = 1.3,
           y = (76-40) /2 + 40 ) +
      annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 44,
           ymax = 98,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = 3,
               label = "Javelin",
           x = 1.3,
           y = (98-44) /2 + 44 ) +
     annotate("rect", 
           xmin = -Inf,
           xmax = Inf,
           ymin = 88,
           ymax = 230,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Thrusting spear",
               size = 3,
           x = 1.3,
           y = (230-88) /2 + 88 ) +
  geom_boxplot(alpha = 0.1) +
  geom_quasirandom(alpha = 0.4,
                   size = 3) +
  theme_bw() +
  labs(title = "TCSA of Korean stemmed points", 
       y = "TCSA")
```
1. The mean TCSA of the Korean stemmed point is `mean(tcsa$min_tcsa)`, and SD TCSA is `sd(tcsa$min_tcsa)`. 
2. According to the Lombard's TCSA range, Korean stemmed points are included mostly in the categories of javelins and thrusting spear tips but overall the boxplot represents wide TCSA variation. 
3. The individual weapon categories of TCSA range are overlapped a lot. We chose the specific colors for the annotation boxes to match with Figure 5 of Lombard's 2021 QSR paper.

```{r}

### create column for "n = xx "
tcsa_raw <- tcsa %>% 
  group_by(raw_materials_top) %>% 
  add_tally() %>% 
  filter(n>1) %>% 
  mutate(label_raw = as.factor(paste0(raw_materials_top,' (N = ', n, ")"))) %>%   
  select(-n) %>%
  nest(-raw_materials_top, -label_raw)

tcsa_site <- tcsa %>% 
  group_by(site_name_top) %>% 
  add_tally() %>% 
  filter(n>1) %>% 
  mutate(label_site = as.factor(paste0(site_name_top,' (N = ', n, ")"))) %>%   
  select(-n) %>%
  nest(-site_name_top, -label_site) %>% 
  select(-data)

tcsa_label <- tcsa %>% 
  left_join(tcsa_raw) %>% 
  left_join(tcsa_site)
```

```{r}
#### histogram for min tcsa to prepare for exploring the data
all <- ggplot(tcsa_label,
              aes(min_tcsa)) +
  geom_histogram(color="black", fill="grey90") +
  theme_bw() +
  labs(x = "TCSA", 
       y = "Count")
```

@@

```{r, fig.height=5}
# Facet by raw materials
all +  facet_wrap(~label_raw,
                  scales = "free_y",
                  ncol = 1)

```

1) About half of the stemmed points made out of shale. 
2) Acidic volcanic rocks tend to have more screwed and lower TCSA value. 
3)

## Size (Weight)
```{r}
# how does the size-tcsa relationship vary by raw material?

# remove NAs
tcsa_mass_no_na <-
  tcsa %>%
 drop_na(`Weight(g)`)

ggplot(tcsa_mass_no_na,
       aes(`Weight(g)`,
           min_tcsa,
           colour = raw_materials_top)) +
  geom_point() +
  theme_bw()
```

@@

```{r}
# Facet by site
all +  facet_wrap(~label_site,
                 scales = "free_y",
                  ncol = 1)
```
